Adaptive Hybrid Screening for Efficient Lasso Optimization

### 2021-09-08 jscs/response/response.tex

\documentclass{article}
\usepackage[left=1in,right=1in,bottom=1in,top=1in]{geometry}
\usepackage{amsmath}

\begin{document}

\subsubsection*{Response to comments by Reviewer 1}

Thank you very much for your detailed reading of the paper and your insightful comments and questions. You raise a number of good points and we have modified the manuscript in several places based upon your feedback.  In our opinion, these revisions have nicely improved upon the original submission.

\begin{enumerate}

\item \emph{Numerical studies only focus on the computational performances of the algorithms. One may be interested in some details in the proposed algorithm: the size of variables remaining from the safe or strong rules, the number of KKT condition checks, the adaptive batch sizes, etc. It would be good to add the numerical studies for details in the algorithm.}

\item \emph{The paper assumes that the predictive variables are standardized for linear and logistic regression models. For group lasso, the additional orthogonal assumption are required for the proposed method. I’m just wondering if the proposed method is applicable without the standardization assumptions. Specifically, is it possible to develop (or exist) the sequential safe and strong rules without the standardization assumptions.}

\item \emph{In page 10, line 13, ``We also assume that post-convergence check will usually pass in the first iterations''. I’m not sure it is reasonable.}

\item \emph{In page 8, line 19, ``the reference solutions are continually updated''. The reference solutions are not always updated. So, I’m not sure the term ``continually'' is appropriate.}

\item \emph{In Figure 3, the green line (HSSR) pattern is quite different with others. Could you explain it?}

\item \emph{In Section 6(Conclusion), there does not exist the sequential safe rules for sparse Cox regression, Poisson regression, and support vector machines. It would be good to explain the existence of the sequential strong rules for each problem.}

\item \emph{In Theorem 4.1, $x_* \in \{ x_j : \hat{\beta}_j(\lambda_l) \ne 0 \}$. Is the $x_*$ arbitrary? It seems that further explanation for $x_*$ is needed.}

\item \emph{Minor corrections:} These have all been made. Thank you very much for pointing them out.

\end{enumerate}

\end{document}

### 2021-09-08 jscs/adaptive-hssr.tex
response: Draft started

This complexity analysis will only consider calculations with $O(np)$ complexity. The remaining calculations, those with $O(p)$ complexity for example, are not only of lower order, but also consume less memory and thus usually do no require out-of-core computing. Calculation of $\{x_j^Ty\}_{j=1}^p$ is needed only once for the whole solution path; we ignore this as well as its computational cost is the same across all approaches. We assume that strong rule screening is always powerful in the sense that its set of non-discarded features, $\mathcal{R}_{l+b}$, satisfies $|\mathcal{R}_{l+b}|=o(p) \approx 0$, but that safe rule screening, with corresponding set $\mathcal{S}_{l+b}$, may not be. We also assume that post-convergence check will usually pass in the first iteration and thus is only needed approximately once per $\lambda$. These two assumptions are supported by experiments from the origin paper of SSR \citep{Tibshirani2012}, many other papers studying feature screening, and our manuscript as discussed in the Appendix \ref{sec:ap}.

@@

\begin{table}[H]
\centering
\begin{tabular}{lllll}
\toprule
 & GENE & MNIST & GWAS & NYT \\
Screening method & n=536 & n=784 & n=313 & n=5,500 \\
       & p=17,322 & p=60,000 & p=660,496 & p=55,000 \\
\midrule
AC & 2.14 (0.03) & 6.82 (0.07) & 46.67 (0.13) & 59.20 (1.89) \\
SSR & 1.34 (0.01) & 5.20 (0.01) & 25.34 (0.08) & 35.40 (0.44) \\
SEDPP & 1.38 (0.01) & 8.40 (0.45) & 25.48 (0.05) & 53.69 (3.81) \\
HSSR & 1.21 (0.01) & 3.44 (0.07) & 23.85 (0.06) & 31.05 (0.80) \\
AHR & \textbf{0.71 (0.01)} & \textbf{1.15 (0.02)} & \textbf{8.80 (0.03)} & \textbf{12.71 (1.05)} \\
\bottomrule
\end{tabular}
\caption{Average computing time in second (standard error) for solving the lasso model}
\label{Tab:real}
\end{table}

AHR again outperforms other screening methods on all data sets, faster than other methods by factor ranging from 70\% to 630\%. Also note that the performance of SEDPP varies considerably for the MNIST and NYT data sets, where the response is randomly selected in each replication, indicating that the speed of SEDPP is sensitive to the data. AHR is built with SEDPP as the safe rule but does not share this issue of instability because it can adaptively adjust itself according the performance of the safe rule. Besides the computation time, we also look at a variaty of other statistics in the experiments, as discussed in the Appendix \ref{sec:ap}.

@@

\appendix
\appendixpage

\section{Supplementary Real Data Analysis Results}
\label{sec:ap}

At the first replication on each of the 4 data sets, we first record the number of features discarded by each screening method at each $\lambda$ value along the path of $L=100$ values. Features discarded by SEDPP are safely discarded and features discarded by SSR are not safely discarded and require additional checks. For the adaptive hybrid rule, we use AHR\_Safe to denote the features that are safely discarded first by adaptive EDPP and AHR\_Strong to denote features that discarded afterward by adaptive SSR and require additional checks. The results are summarized in Figure \ref{fig:ap1}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.82\textwidth]{app1.eps}    \caption{Comparing number of discards by screening methods for lasso model along $\lambda$ values.}
    \label{fig:ap1}
\end{figure}

Note each jump of the number of discards for AHR method indicates an update of reference because a newer reference leads to higher screening power. Our adaptive updating algorithm determines the timing of update dynamically depending on the screening performance. For all four data sets, with about only 10 updates, the adaptive hybrid screening can safely discard a great number of features at most $\lambda$ values.

Next we compare the number of rounds of KKT conditions checks and the number of KKT conditions checks on individual features, needed for AHR and SSR respectively, as shown in Table \ref{Tab:kkt1} and Table \ref{Tab:kkt2}. In a round of KKT conditions check, all features that are not discarded by the safe rule are checked, and a round a KKT conditions check potentially include a large number of KKT conditions checks on individual features.

\begin{table}[H]
\centering
\begin{tabular}{lllll}
\toprule
Screening method & GENE & MNIST & GWAS & NYT \\
\midrule
SSR & 99 & 99 & 99 & 99 \\
AHR & 99 & 99 & 99 & 99 \\
\bottomrule
\end{tabular}
\caption{Number of rounds of KKT conditions checks required for the whole path}
\label{Tab:kkt1}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{lllll}
\toprule
Screening method & GENE & MNIST & GWAS & NYT \\
\midrule
SSR & 1,712 & 5,943 & 60,986 & 5,429 \\
AHR & 172 & 286 & 2,353 & 622 \\
\bottomrule
\end{tabular}
\caption{Number of KKT conditions checks on individual features ($\times10^3$) required for the whole path}
\label{Tab:kkt2}
\end{table}

Although a path include $L=100$ $\lambda$ values, the solution at the first value $\lambda_{0}$ is trivially $\hat{\beta}(\lambda_0)=0$ so we are essentially solving along $99$ $\lam$ values. If the strong rule does not make a violation, the minimal number of rounds of KKT conditions checks required will be $99$. In all four data sets, neither SSR nor the adaptive SSR in AHR ever makes a violation. However two methods are very different in terms of the number of KKT conditions checks on individual features, which proportionally increases the computational cost. AHR requires much less checks on individual features because the adaptive EDPP narrows down the set of features that need to be checked by a magnitude.

>>>>

This complexity analysis will only consider calculations with $O(np)$ complexity. The remaining calculations, those with $O(p)$ complexity for example, are not only of lower order, but also consume less memory and thus usually do no require out-of-core computing. Calculation of $\{x_j^Ty\}_{j=1}^p$ is needed only once for the whole solution path; we ignore this as well as its computational cost is the same across all approaches. We assume that strong rule screening is always powerful in the sense that its set of non-discarded features, $\mathcal{R}_{l+b}$, satisfies $|\mathcal{R}_{l+b}|=o(p) \approx 0$, but that safe rule screening, with corresponding set $\mathcal{S}_{l+b}$, may not be. We also assume that post-convergence check will usually pass in the first iteration and thus is only needed approximately once per $\lambda$. These two assumptions are supported by experiments from the origin paper of SSR \citep{Tibshirani2012}, many other papers studying feature screening, and our manuscript as discussed in the Appendix.

@@

\begin{table}[ht]
\centering
\begin{tabular}{lllll}
\toprule
 & GENE & MNIST & GWAS & NYT \\
Screening method & n=536 & n=784 & n=313 & n=5,500 \\
       & p=17,322 & p=60,000 & p=660,496 & p=55,000 \\
\midrule
AC & 2.14 (0.03) & 6.82 (0.07) & 46.67 (0.13) & 59.20 (1.89) \\
SSR & 1.34 (0.01) & 5.20 (0.01) & 25.34 (0.08) & 35.40 (0.44) \\
SEDPP & 1.38 (0.01) & 8.40 (0.45) & 25.48 (0.05) & 53.69 (3.81) \\
HSSR & 1.21 (0.01) & 3.44 (0.07) & 23.85 (0.06) & 31.05 (0.80) \\
AHR & \textbf{0.71 (0.01)} & \textbf{1.15 (0.02)} & \textbf{8.80 (0.03)} & \textbf{12.71 (1.05)} \\
\bottomrule
\end{tabular}
\caption{Average computing time in seconds (standard error) for solving the lasso model}
\label{Tab:real}
\end{table}

AHR again outperforms other screening methods on all data sets, faster than other methods by factor ranging from 70\% to 630\%. Also note that the performance of SEDPP varies considerably for the MNIST and NYT data sets, where the response is randomly selected in each replication, indicating that the speed of SEDPP is sensitive to the data. AHR is built with SEDPP as the safe rule but does not share this issue of instability because it can adaptively adjust itself according the performance of the safe rule. Besides the computation time, we also look at a variaty of other statistics in the experiments, as discussed in the Appendix.

@@
